Sakura Blossom Prediction Model for Japan

### 2024-11-27 paper/paper.qmd
first draft!

---
title: "Sakura Blossom Prediction Model for Japan"
subtitle: "Forecasting Sakura Blossom Using Bayesian Spline Regression"
author: 
  - Shanjie Jiao
thanks: "Code and data are available at: https://github.com/Jie-jiao05/Sakura-Blossom-Prediction-Model."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib

@@

library(tidyverse)
library(dplyr)
library(here)
library(readr)
library(lubridate)
library(rvest)
library(palmerpenguins)
library(maps)
library(ggplot2)
library(knitr)
library(arrow)
library(rstanarm)

# Define paths to data and model
analysis_modern_sakura_data <- here("data", "02-analysis_data", "analysis_modern_sakura_data.parquet")
temperature_data <- here("data", "02-analysis_data", "temperature_data.csv")

# Load the data
sakura_data <- read_parquet(analysis_modern_sakura_data)  # Load Parquet data
tem_data <- read_csv(temperature_data)                    # Load CSV data

@@

train_summary <- data.frame(
  Statistic = c("Unique Stations", "Mean Flower Time", "Mean Temperature", "Mean Full Blossom"),
  Value = c(
    length(unique(train$station_id)),
    mean(train$flower_day_in_numeric, na.rm = TRUE),
    mean(train$month_mean_temp, na.rm = TRUE),
    mean(train$full_bloom_day_in_numeric, na.rm = TRUE)
  ),
  Dataset = "Train Data"
)

# Combine summaries
summary_data <- rbind(test_summary, train_summary)

# Create sakura-themed visualization with transparency
ggplot(summary_data, aes(x = Statistic, y = Value, fill = Dataset)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, alpha = 0.7) + # Add transparency with alpha
  geom_text(
    aes(label = round(Value, 1)), 
    position = position_dodge(width = 0.7), 
    vjust = -0.2, 
    size = 2.5,
    color = "black"
  ) +
  scale_fill_manual(
    values = c("Test Data" = rgb(255, 193, 204, maxColorValue = 255, alpha = 180),  # Sakura pink with transparency
               "Train Data" = rgb(163, 214, 167, maxColorValue = 255, alpha = 180)) # Soft green with transparency
  ) +
  labs(
    title = "Comparison of Train and Test Data",
    x = "Statistic",
    y = "Value"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16, color = "#6A1B9A"),
    plot.subtitle = element_text(hjust = 0.5, size = 12, color = "#6A1B9A"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    legend.title = element_blank(),
    legend.position = "top"
  )
```

@@

## Result of the Prediction Model

@@

```{r}
#| label: fig-obs   # For citation in Quarto
#| tbl-cap: "Testing Data: Observed vs Predicted Flowering Days"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
# === Step 7: Visualize Results (Optional) ===
# Scatter plot: Observed vs. Predicted (Training Data)
ggplot(train, aes(x = flower_day_in_numeric, y = predicted_flower_day)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Observed Flowering Day (Train)",
    y = "Predicted Flowering Day (Train)",
    title = "Training Data: Observed vs Predicted Flowering Days"
  ) +
  theme_minimal()

# Scatter plot: Observed vs. Predicted (Testing Data)
ggplot(test, aes(x = flower_day_in_numeric, y = predicted_flower_day)) +
  geom_point(color = "green", alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Observed Flowering Day (Test)",
    y = "Predicted Flowering Day (Test)",
    title = "Testing Data: Observed vs Predicted Flowering Days"
  ) +
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

#library(rstanarm)
#library(modelsummary)

#first_model <-
#  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

#modelsummary::modelsummary(
#  list(
#    "First model" = first_model
#  ),
#  statistic = "mad",
#  fmt = 2
#)
```

@@

## First discussion point {#sec-first-point}

@@

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

#pp_check(first_model) +
#  theme_classic() +
#  theme(legend.position = "bottom")

#posterior_vs_prior(first_model) +
#  theme_minimal() +
#  scale_color_brewer(palette = "Set1") +
#  theme(legend.position = "bottom") +
#  coord_flip()
```

@@

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```

>>>>

---
title: "Sakura Blossom Prediction Model for Japan"
subtitle: "Forecasting Sakura Blossom Using Bayesian Spline Regression"
author: 
  - Shanjie Jiao
thanks: "Code and data are available at: https://github.com/Jie-jiao05/Sakura-Blossom-Prediction-Model."
date: today
date-format: long
abstract: "This study develops a Bayesian hierarchical linear regression model to predict sakura blooming dates across Japan by analyzing key predictors such as temperature, latitude, and longitude. The model demonstrates strong predictive accuracy, capturing over 95% of the variability in flowering dates and revealing clear regional and climatic patterns. The findings provide valuable insights into the effects of climate change on sakura phenology, aiding tourism planning and ecological conservation. Future improvements could incorporate dynamic modeling techniques, additional environmental variables, and broader datasets to enhance the model's precision and adaptability."
format: pdf
toc: true
number-sections: true
bibliography: references.bib

@@

library(tidyverse)
library(dplyr)
library(here)
library(readr)
library(lubridate)
library(rvest)
library(palmerpenguins)
library(maps)
library(ggplot2)
library(knitr)
library(arrow)
library(rstanarm)
library(knitr)
library(plotly)

# Define paths to data and model
analysis_modern_sakura_data <- here("data", "02-analysis_data", "analysis_modern_sakura_data.parquet")
temperature_data <- here("data", "02-analysis_data", "temperature_data.csv")
history_sakura <- here("data", "02-analysis_data", "analysis_historical_sakura_data.parquet")

# Load the data
sakura_data <- read_parquet(analysis_modern_sakura_data)  # Load Parquet data
tem_data <- read_csv(temperature_data)                    # Load CSV data
history_data <- read_parquet(history_sakura)

@@

train_summary <- data.frame(
  Statistic = c("Unique Stations", "Mean Flower Time", "Mean Temperature", "Mean Full Blossom"),
  Value = c(
    length(unique(train$station_id)),
    mean(train$flower_day_in_numeric, na.rm = TRUE),
    mean(train$month_mean_temp, na.rm = TRUE),
    mean(train$full_bloom_day_in_numeric, na.rm = TRUE)
  ),
  Dataset = "Train Data"
)
# Combine summaries
summary_data <- rbind(test_summary, train_summary)

# Create sakura-themed visualization with smaller text sizes
ggplot(summary_data, aes(x = Statistic, y = Value, fill = Dataset)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, alpha = 0.7) + # Add transparency with alpha
  geom_text(
    aes(label = round(Value, 1)), 
    position = position_dodge(width = 0.7), 
    vjust = -0.2, 
    size = 2,  # Reduce text size for labels
    color = "black"
  ) +
  scale_fill_manual(
    values = c("Test Data" = rgb(255, 193, 204, maxColorValue = 255, alpha = 180),  # Sakura pink with transparency
               "Train Data" = rgb(163, 214, 167, maxColorValue = 255, alpha = 180)) # Soft green with transparency
  ) +
  labs(
    title = "Comparison of Train and Test Data",
    x = "Statistic",
    y = "Value"
  ) +
  theme_minimal(base_size = 10) +  # Reduce base font size
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12, color = "black"), # Smaller title size
    axis.text.x = element_text(angle = 43, hjust = 1, size = 9), # Smaller x-axis labels
    axis.text.y = element_text(size = 9),  # Smaller y-axis labels
    legend.title = element_blank(),
    legend.position = "top",
    legend.text = element_text(size = 9)  # Smaller legend text
  )

@@

## Result of the Analysis Data
In @fig-rela_result is the result of correlation between each predictor variable, including latitude, longitude, and monthly mean temperature, and flowering day.
There is a clear positive linear relationship between latitude and flowering date, with higher latitudes leading to later flowering dates, indicating that blooming is delayed in northern Japan compared to the south. 
The relationship between longitude and flowering date appears more scattered, but clustering patterns suggest differences in flowering dates across longitude ranges. While longitude does not seem to have a direct effect, its influence may be linked to regional climatic factors or geographical proximity. 
Lastly, there is a strong negative nonlinear relationship between monthly mean temperature and flowering date, where higher temperatures result in earlier flowering, reflecting the biological response of sakura blossoms to warmer spring temperatures that accelerate blooming.

```{r}
#| label: fig-rela_result
#| fig-cap: "Relationships Between Predictors and Flowering Day"
#| echo: false
#| warning: false
#| message: false
# Load required libraries
library(ggplot2)
library(tidyr)

# Reshape the data to long format for faceting
sakura_long <- sakura_data %>%
  pivot_longer(cols = c(month_mean_temp, latitude, longitude),
               names_to = "Predictor", values_to = "Value")

# Create a faceted scatter plot with a professional red tone
facet_plot <- ggplot(sakura_long, aes(x = Value, y = flower_day_in_numeric)) +
  geom_point(alpha = 0.8, size = 1, color = "#FF9999") +  # Professional red tone
  facet_wrap(~ Predictor, scales = "free_x", ncol = 3) +    # Arrange in a single row
  labs(
    title = "Relationships Between Predictors and Flowering Day",
    x = "Predictor Value",
    y = "Flowering Day (Numeric)"
  ) +
  theme_minimal(base_size = 12) +  # Adjust font size for professional look
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold", color = "black"),  # Deep red for title
    strip.text = element_text(size = 12, face = "bold", color = "#800000"),  # Bold red facet labels
    axis.title = element_text(size = 11, color = "#4D4D4D"),  # Neutral gray for axis titles
    axis.text = element_text(size = 10, color = "#4D4D4D"),  # Neutral gray for axis text
    panel.grid.major = element_line(color = "#EAEDED"),  # Subtle light gray grid lines
    panel.grid.minor = element_blank(),  # No minor grid lines
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +  # Add spacing for clarity on the y-axis
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1)))    # Add spacing for clarity on the x-axis

# Display the plot
print(facet_plot)

## Result of the Prediction Model
### Performance Overview Analysis

The scatter plot @fig-pred illustrates a strong alignment between observed and predicted sakura flowering dates for both training and testing datasets. Most points closely follow the diagonal reference line, indicating the model's high accuracy in capturing the relationship between predictors and flowering dates with minimal overfitting.

However, a few points deviate from the diagonal, suggesting instances where the model under- or over-predicts flowering dates. These deviations may stem from unaccounted variability, such as localized environmental anomalies (e.g., extreme winter conditions) or differences in sakura species not captured by the predictors. Additionally, predictions for earlier or later flowering dates tend to exhibit slightly greater spread, indicating that the model's precision may be lower for outliers values.

```{r}
#| label: fig-pred
#| fig-cap: "Observed vs. Predicted Flowering Dates"
#| echo: false
#| warning: false
#| message: false
# === Step 3: Generate Predictions ===
# Training predictions
train_predictions <- posterior_predict(flower_model)
train$predicted_flower_day <- colMeans(train_predictions)

# Combine metrics into a table
metrics_table <- data.frame(
  Metric = c("RMSE", "MAE", "R^2"),
  Training = train_metrics,
  Testing = test_metrics
)
# Prepare Data
train$Set <- "Train"
test$Set <- "Test"
combined_data <- rbind(train, test)

# Ensure 'Set' is a valid factor
combined_data$Set <- as.factor(combined_data$Set)
combined_data$Set <- droplevels(combined_data$Set)

ggplot(data = combined_data, aes(
  x = flower_day_in_numeric, 
  y = predicted_flower_day, 
  color = Set
)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Observed vs. Predicted Flowering Dates",
    x = "Observed Flowering Dates",
    y = "Predicted Flowering Dates",
    color = "Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

@@

@@

## Interpretation and Implications
Temperature emerged as the most significant predictor, with higher monthly mean temperatures accelerating flowering dates. This aligns with established ecological theories linking warming trends to phenological shifts in plant species. Latitude also showed a strong positive effect, reflecting the gradient of delayed flowering in colder, northern regions compared to southern ones. Interestingly, longitude had a less pronounced effect, suggesting that the east-west climatic variations within Japan are relatively minor compared to the north-south differences. These findings emphasize the importance of geographic and climatic factors in shaping flowering patterns and suggest that warming trends will continue to drive earlier sakura blooming in future decades.

## Limitation
The model demonstrates strong predictive performance but has several limitations that should be addressed. By assuming that all regions respond uniformly to climate variables, the dataset may oversimplify the complex ecological processes influencing cherry blossom flowering. The important factors such as soil moisture variations, urban heat effects, and species-specific traits are not explicitly accounted for, potentially overlooking significant regional nuances. Furthermore, the absence of geographic predictors like elevation and detailed precipitation data likely contributes to larger prediction errors for outliers, particularly in regions with extreme or atypical flowering conditions. Additionally, the dataset does not differentiate between sakura varieties, despite substantial variation in flowering times driven by species-specific biological and environmental factors. This lack of certain specices may reduce the model’s precision, especially in areas with diverse sakura species and environmental contexts.

@@

## Implications of the Sakura Prediction Model for Climate Sensitivity and Tourism Planning

The establishment of this prediction model will significantly enhance understanding of sakura’s sensitivity to climate, as well as inform tourism and cultural planning associated with sakura. Due to the influence of latitude and temperature, sakura blooms at different times across Japan, with warmer southern regions blooming earlier than cooler northern areas. These regional differences in flowering dates offer an opportunity to better distribute tourist influx in densely visited areas, alleviating pressure on urban infrastructure and facilities. Accurate predictions of flowering times can lead to more effective resource allocation, enabling tourists to plan their trips to Japan at the optimal time. This allows for a more tailored travel experience, with visitors enjoying sakura in different regions at different times, while simultaneously boosting regional economic development. Tourism bureaus can utilize these predictions to adjust event schedules, aligning activities with the expected bloom dates to provide an optimal visitor experience. Additionally, regional collaboration could encourage staggered travel itineraries, where tourists follow the sakura bloom progression from south to north, effectively dispersing demand across time and space. Northern regions with later blooming may benefit from an extended tourist season, whereas central regions such as Kyoto might face challenges with overcrowding if their peak bloom overlaps with other popular destinations.

Beyond its primary function of predicting flowering times, the model incorporates precise factors such as temperature and geographic location, offering valuable insights into the ecological dynamics of sakura. The marked sensitivity of flowering dates to temperature underscores the characteristic responses of plant phenology to climate change, often referred to as a "natural barometer." By systematically analyzing the relationships between climate factors and flowering dates, the model reveals how sakura responds to seasonal temperature fluctuations, such as earlier blooming due to rising temperatures, as well as regional adaptations influenced by latitude. Furthermore, this model provides a powerful tool for evaluating the long-term impacts of climate change. For example, by simulating future climate scenarios under varying greenhouse gas emission trajectories, the model can predict shifts in sakura phenology over the coming decades. This is essential for understanding the broader implications of climate change on plant phenology, both in Japan and globally. Varieties with heightened sensitivity to temperature changes may face greater survival pressures as climate change intensifies, and insights from the model can guide conservation strategies, helping to identify priority regions and species that require focused protection efforts.

## Future Directions for Improving the Sakura Prediction Model

For a more precise prediction, several improvements should be implemented. First, incorporating additional predictors such as elevation and microclimate factors would enhance accuracy by accounting for localized influences on blooming, particularly in regions with diverse topography or urban heat effects. Including soil conditions and precipitation data would provide a deeper understanding of environmental factors affecting sakura phenology, especially in areas vulnerable to drought or irregular rainfall. Second, expanding the spatial and temporal coverage of the dataset is crucial. Integrating records from other sakura-growing regions, such as South Korea, China, and the United States, would enable cross-regional comparisons and distinguish universal trends from region-specific climate responses. Extending the dataset with more historical and continuously updated records would also facilitate a thorough analysis of long-term flowering trends. Lastly, adopting dynamic or time-series modeling techniques would improve the model’s ability to capture year-to-year variability and respond to sudden climate anomalies, such as extreme weather events. Methods like dynamic Bayesian networks or ARIMA models could enhance predictions in increasingly erratic climatic conditions. Collectively, these advancements would make the model more accurate, flexible, and insightful for understanding sakura phenology and its interactions with climate change.
\newpage

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. 
This shows the density overlay plot demonstrates a strong alignment between the observed sakura flowering dates and the model's posterior predictions. The observed density, represented by the dark blue line, closely matches the predictive densities (lighter blue lines) across most of the distribution, indicating that the model captures the central tendency and major patterns effectively. The primary peak around 80 is well-represented, showing the model's strength in predicting the most common flowering dates. Additionally, the secondary peak around 120 aligns reasonably well, suggesting the model's ability to handle more complex patterns, such as bimodal distributions. However, slight discrepancies are observed in the tails of the distribution, particularly below 40 and above 140, where the predictive densities deviate from the observed data. These divergences indicate potential challenges in capturing extreme values or outliers, which may be addressed by including additional predictors, such as microclimatic conditions or species-specific traits.

In @fig-ppcheckandposteriorvsprior we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
#| warning: false 
#| message: false
#| fig-align: center

# Generate posterior predictive samples for the training data
posterior_samples <- posterior_predict(flower_model)

# Plot posterior predictive distribution
library(bayesplot)
ppc_dens_overlay(
  y = train$flower_day_in_numeric, 
  yrep = posterior_samples[1:100, ] # Use 100 draws for visualization
) +
  ggtitle("Posterior Predictive Check: Density Overlay") +
  theme_minimal()

@@

@fig-mcmc-convergence
The MCMC diagnostics provide strong evidence of convergence and reliability in the posterior estimates. The trace plots for key parameters, including latitude, longitude, and month_mean_temp, demonstrate good mixing across all four chains, with no discernible trends or drift over iterations. Each parameter fluctuates within a stable range, reflecting consistent exploration of the posterior distribution. For instance, latitude ranges between approximately 4.5 and 5.6, while longitude and month_mean_temp remain similarly stable. The R-hat diagnostics further confirm convergence, with all values close to 1.0, indicating that the chains are well-mixed and have reached equilibrium. These results collectively validate the robustness of the MCMC algorithm, ensuring that the posterior estimates accurately reflect the relationships between the predictors and sakura flowering dates. This provides a solid foundation for subsequent inferences and model interpretation.

```{r}
#| label: fig-mcmc-convergence
#| fig-cap: "MCMC Convergence Diagnostics"
#| fig-subcap: ["Trace Plot for Key Parameters", "R-hat Diagnostics"]
#| echo: false
#| warning: false 
#| message: false
#| fig-align: center

library(bayesplot)
library(gridExtra)

# Extract posterior samples from the model
posterior_samples <- as.array(flower_model)

# === Trace Plot ===
# Include all key parameters: latitude, longitude, and month_mean_temp
key_parameters <- c("latitude", "longitude", "month_mean_temp")
trace_plot <- mcmc_trace(
  posterior_samples,
  pars = key_parameters,
  facet_args = list(ncol = 1)
) +
  ggtitle("Trace Plot for Key Parameters") +
  theme_minimal()

# === Filtered R-hat Diagnostics ===
# Calculate R-hat values for all parameters
rhat_values <- rhat(flower_model)

# Include R-hat values for key parameters
selected_parameters <- c("latitude", "longitude", "month_mean_temp", "sigma") # Add "longitude" here
filtered_rhat_values <- rhat_values[selected_parameters]

# Plot R-hat diagnostics for filtered parameters
rhat_plot <- mcmc_rhat(filtered_rhat_values) +
  ggtitle("Filtered R-hat Diagnostics") +
  theme_minimal()

# === Combine Plots ===
grid.arrange(trace_plot, rhat_plot, ncol = 2)
